---
title: 'Neurosurgical Patents: Physicians vs. Non-Physicians'
author: "Ryan G. Chiu"
date: "5/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
#Import libraries
library(foreign)
library(readxl)
library(dplyr)
library(MASS)
library(MatchIt)
library(RColorBrewer)
library(knitr)
library(tableone)
library(Matching)
library(pastecs)
library(stringr)
library(epitools)
library(elixhauser)
library(jsonlite)
library(rvest)
library(RCurl)
```

``` {r, echo=FALSE}
# Function definitions:
get_stats <- function(vals) {
  good_ones <- vals[!is.na(vals)]
  return(c(length(good_ones), mean(good_ones), sd(good_ones)))
}

# Substitute non-English characters with equivalent English ones (NPI database won't support otherwise)
to.plain <- function(s) {

   # 1 character substitutions
   old1 <- "šžþàáâãäåçèéêëìíîïðñòóôõöùúûüýŠŽÞÀÁÂÃÄÅÇÈÉÊËÌÍÎÏÐÑÒÓÔÕÖÙÚÛÜÝ"
   new1 <- "szyaaaaaaceeeeiiiidnooooouuuuySZYAAAAAACEEEEIIIIDNOOOOOUUUUY"
   s1 <- chartr(old1, new1, s)

   # 2 character substitutions
   old2 <- c("œ", "ß", "æ", "ø")
   new2 <- c("oe", "ss", "ae", "oe")
   s2 <- s1
   for(i in seq_along(old2)) s2 <- gsub(old2[i], new2[i], s2, fixed = TRUE)

   s2
}

# Extract number of citations given HTML of patent full-text webpage
get_citations_from_html <- function(html) {
  h2_vector <- html %>% html_nodes('h2') %>% html_text() %>% str_trim() %>% unlist()
  citation_text_matches <- grep('Cited By \\(', h2_vector, value=TRUE)
  
  # Can't be too safe with this stuff -- always look for possibility of null values ;)
  citation_text <- ifelse(length(citation_text_matches) > 0, citation_text_matches[1], 'Cited By (0)')
  
  # Now, let's get the number from the text ('Cited By (the_number_we_want)')
  number_of_citations <- as.numeric(gsub("\\D", "", citation_text))
  
  number_of_citations
}
```

Loading Google Patents output file ...
```{r, echo=FALSE}
# Load patents file
setwd("/Users/Ryan/Downloads/")
data <- read.csv("data22.csv", header=TRUE)
```

Convert inventors string to array of names, using comma as delimiter:
```{r, echo=FALSE}
inventors_list <- as.vector(data$inventors)
i <- 0
for (inventors in inventors_list) {
  inventor_array <- str_split(inventors, ", ")
  inventors_list[i] <- inventor_array
  i <- i + 1
}
```

We should now have a vector of vectors (array of arrays) for the inventors of each patents

Now, we have to iterate through this updated vector and run each set of inventors through the NPI JSON database to identify neurosurgeon inventor presence. We will essentially need a nested for loop for this: 
```{r}
# This is the vector for our independent variable (i.e. is there a physician inventor aboard?) that will be applied as a column in our final dataframe
is_neurosurgeon <- c()
index <- 0

for (inventors in inventors_list) {
  physician_count <- 0
  for (inventor in inventors) {
    name <- str_split(inventor, " ")
    json_string <- paste0("https://npiregistry.cms.hhs.gov/api/?version=2.1&taxonomy_description=Neurological%20Surgery&first_name=", to.plain(name[[1]][1]), "&last_name=", to.plain(name[[1]][length(name[[1]])]))

    attempt <- 0
    json <- NULL
    while( is.null(json) && attempt <= 100) {
      attempt <- attempt + 1
      try(
        json <- fromJSON(json_string)
      )
    } 
    
    if (!is.null(json$result_count) && json$result_count > 0) {
      physician_count <- physician_count + 1
    } else {
      physician_count <- physician_count
    }
  }
  
  index <- index + 1
  print(sprintf("Identifying neurosurgeons: %d of %d dataset rows processed ...", index, nrow(data)-1))
  is_neurosurgeon <- c(is_neurosurgeon, physician_count)
}
```

Okay, phew -- alright one more time, this time to identify the ORTHO BROS:
```{r}
is_orthopod <- c()
index <- 0

for (inventors in inventors_list) {
  physician_count <- 0
  for (inventor in inventors) {
    name <- str_split(inventor, " ")
    json_string <- paste0("https://npiregistry.cms.hhs.gov/api/?version=2.1&taxonomy_description=Orthopaedic%20Surgery&first_name=", to.plain(name[[1]][1]), "&last_name=", to.plain(name[[1]][length(name[[1]])]))
    
    attempt <- 0
    json <- NULL
    while( is.null(json) && attempt <= 100) {
      attempt <- attempt + 1
      try(
        json <- fromJSON(json_string)
      )
    } 
    
    if (!is.null(json$result_count) && json$result_count > 0) {
      physician_count <- physician_count + 1
    } else {
      physician_count <- physician_count
    }
  }
  
  index <- index + 1
  print(sprintf("Identifying orthopaedic surgeons: %d of %d dataset rows processed ...", index, nrow(data)-1))
  is_orthopod <- c(is_orthopod, physician_count)
}
```

Okay, now let's scrape the full-text webpages (HTML) for number of citations. There should be a header called "Cited By" followed by a number in parentheses for granted patents

HTML:
<h2>Cited By (##)</h2>
```{r}
# Change this to data 1,2,3,4,5
links <- as.vector(data$result_link)

numbers_of_citations <- c()
index <- 0

for (link in links) {
  n <- NULL
  attempt <- 0
  while( is.null(n) && attempt <= 100) {
    attempt <- attempt + 1
    try(
      n <- get_citations_from_html(read_html(link))
    )
  } 
  
  numbers_of_citations <- c(numbers_of_citations, n)
  index <- index + 1
  
  # ... and replace here too
  print(sprintf("Retrieving number of citations: %d of %d dataset rows processed ...", index, nrow(data)-1))
}
```

Now that we have bool vectors of calculated columns, let's append it onto our dataset as the last column:
```{r}
final_data <- cbind(data, data.frame(is_neurosurgeon), data.frame(is_orthopod), data.frame(numbers_of_citations))
```

Okay, now we can do data clean-up before starting stats:
- is_neurosurgeon and is_orthopod are integer arrays -- 0 = no physician, 1+ = physician inventor present --> convert to boolean
- in the data file, grant_date is blank if the patent wasn't granted --> let's make separate variable for this
- application date and grant_date are formatted MM/DD/YY --> let's calculate new variable of "days till patent grant"
- see if assignee is corporation vs. individual by business abbreviations such as "Inc" or "LLC" or "GmbH"
```{r}
final_data <- mutate(final_data, neurosurgeon_involved = ifelse(is_neurosurgeon > 0, "yes", "no"))
final_data <- mutate(final_data, ortho_involved = ifelse(is_orthopod > 0, "yes", "no"))

final_data <- mutate(final_data, granted = ifelse(nchar(as.character(grant_date)) > 0, "yes", "no"))

final_data <- mutate(final_data, days_to_grant = ifelse(granted == "yes", as.Date(as.character(grant_date), format="%m/%d/%Y")-as.Date(as.character(filing_creation_date), format="%m/%d/%Y"), NA))

final_data <- mutate(final_data, assignee_type = grepl("llc",tolower(assignee)) | grepl("lls",tolower(assignee)) | grepl("llp",tolower(assignee)) | grepl("inc",tolower(assignee)) | grepl("gmbh",tolower(assignee)) | grepl("corp",tolower(assignee)) | grepl("co.",tolower(assignee)))
final_data <- mutate(final_data, assignee_type = ifelse(assignee_type, "corporation", "individual"))
```

Now, FINALLY we have all the variables we needed to run stats --> let's export to a new, clean CSV file
``` {r}
write.csv(final_data, 'patent_data_clean.csv')
```